---
title: "R Notebook"
output: html_notebook
---


```{r}
library(dplyr)
library(caret)
library(ggplot2)


Bank <- read.csv("UniversalBank.csv")
##Bank <-UniversalBank
head(Bank)
str(Bank)
Bank1 <- Bank[,c(-1, -5)] ##excluding 2 coloumn form data "ID" and "Zip Code"

##converting Personal loan to factor
Bank1$PersonalLoan=as.factor(Bank1$PersonalLoan)
Bank1$Education = as.factor(Bank1$Education)
levels(Bank1$Education)

dummy_model<-dummyVars(~Education,data=Bank1)
head(predict(dummy_model, Bank1))


Bank1<-cbind(Bank1, predict(dummy_model, Bank1))
Bank1$Education<- NULL
Bank1
View(Bank1)
```


Qs 1) 
Consider the following customer: Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1, and Credit Card = 1. Perform a k-NN classification with all predictors except ID and ZIP code using k = 1. Remember to transform categorical predictors with more than two categories into dummy variables first. Specify the success class as 1 (loan acceptance), and use the default cutoff value of 0.5. How would this customer be classified?
```{r}
set.seed(25)

Test_Data<-read.csv("test.csv")


Test_Data <- Test_Data[,c(-1, -5)] ##excluding 2 coloumn form data "ID" and "Zip Code"

Test_Data$Education.1 = as.factor(Test_Data$Education.1)
Test_Data$Education.2 = as.factor(Test_Data$Education.2)
Test_Data$Education.3 = as.factor(Test_Data$Education.3)
levels(Test_Data$Education)
Test_Data
 
```


```{r}
## data Partioning into training and validation data
Train_Index = createDataPartition(Bank1$PersonalLoan,p=0.6,list=FALSE) 
Train_Data = Bank1[Train_Index,]
Validation_Data = Bank1[-Train_Index,]

summary(Train_Data)
NROW(Train_Data)
NROW(Test_Data)
NROW(Validation_Data)
```


```{r}
## Normalization
# Copy the original data
train.norm.df <- Train_Data
valid.norm.df <- Validation_Data
test.norm.df <- Test_Data
norm.values <- preProcess(Train_Data[, 1:14], method=c("center", "scale"))

train.norm.df[, 1:14] <- predict(norm.values, Train_Data[, 1:14]) 
valid.norm.df[, 1:14] <- predict(norm.values, Validation_Data[, 1:14])
test.norm.df[, 1:13] <- predict(norm.values, Test_Data[, 1:13])
norm.values <- preProcess(Train_Data[, 1:14], method=c("center", "scale"))
```


```{r}
summary(train.norm.df)
```


```{r}
summary(valid.norm.df)
```


```{r}
## Data Plotting

library(ggplot2)
ggplot(Train_Data, aes(x=Age,y=Income, Color="PersonalLoan")) +
  geom_point() 
ggplot(Train_Data, aes(x=Income,y=Experience, Color="Mortagage")) +
  geom_point()
```

Qs 2)
What is a choice of k that balances between overfitting and ignoring the predictor information?

We will now run our model and test on the validation set
```{r}
##Modelling knn
library(FNN)
nn <- knn(train = train.norm.df[, 1:6], test = test.norm.df[, 1:6], 
          cl = train.norm.df[, 7], k = 1, prob=TRUE)
row.names(Train_Data)[attr(nn, "nn.index")]



```

## knn of test set is 1 that is loan would be accept

```{r}
library(caret)
accuracy.df <- data.frame(k = seq(1, 14, 1), accuracy = rep(0, 14))

# compute knn for different k on validation.
for(i in 2:14) {
  knn.pred <- knn(train.norm.df[,-7], valid.norm.df[,-7], 
                  cl = train.norm.df[,7], k = i)
  accuracy.df[i, 2] <- confusionMatrix(knn.pred, valid.norm.df[,7])$overall[1] 
}
accuracy.df
```


Qs 3) 
Show the confusion matrix for the validation data that results from using the best k.
```{r}
 knn.pred <- knn(train.norm.df[,-7], valid.norm.df[,-7], 
                  cl = train.norm.df[,7], k = 3)
  accuracy.df <- confusionMatrix(knn.pred, valid.norm.df[,7])$overall[1]
  accuracy.df
```

```{r}
 CrossTable(x=valid.norm.df[,7],y=knn.pred)
```

Qs 4)
Consider the following customer: Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit Card = 1. Classify the customer using the best k.
```{r}

library(FNN)
nn <- knn(train = train.norm.df[, 1:6], test = test.norm.df[, 1:6], 
          cl = train.norm.df[, 7], k = 3, prob=TRUE)
row.names(train.norm.df)[attr(nn, "nn.index")]
```



Qs 5)
Repartition the data, this time into training, validation, and test sets (50% : 30% : 20%). Apply the k-NN method with the k chosen above. Compare the confusion matrix of the test set with that of the training and validation sets. Comment on the differences and their reason.
```{r}
## repartion the data  into training, validation, and test sets (50% : 30% : 20%)
Bank2 <- read.csv("UniversalBank.csv")
Bank2 <- Bank[,c(-1, -5)] ##excluding 2 coloumn form data "ID" and "Zip Code"
```


```{r}
##converting Personal loan to factor
Bank2$PersonalLoan=as.factor(Bank2$PersonalLoan)
Bank2$Education = as.factor(Bank2$Education)
levels(Bank2$Education)

dummy_model<-dummyVars(~Education,data=Bank2)
head(predict(dummy_model, Bank2))


Bank2<-cbind(Bank2, predict(dummy_model, Bank2))
Bank2$Education<- NULL
Bank2
```


```{r}

##Repartition the data, tnto training, validation, and test sets (50% : 30% : 20%).
set.seed(20)
Train_Index = createDataPartition(Bank2$PersonalLoan,p=0.5, list=FALSE) 
Train_Data = Bank2[Train_Index,]
ValTest_data = Bank2[-Train_Index,]

Validation_Index <- createDataPartition(ValTest_data$PersonalLoan,p=0.6, list=FALSE) 
Validation_Data <- ValTest_data[Validation_Index,]
Test_Data <- ValTest_data[-Validation_Index,]
summary(Train_Data)

```


```{r}
summary(Validation_Data)
```

```{r}
## Normalization
# Copy the original data
train.norm.df1 <- Train_Data
valid.norm.df1 <- ValTest_data
test.norm.df1 <- Test_Data
norm.values <- preProcess(Train_Data[, 1:14], method=c("center", "scale"))

train.norm.df1[, 1:14] <- predict(norm.values, Train_Data[, 1:14]) 
valid.norm.df1[, 1:14] <- predict(norm.values, ValTest_data[, 1:14])
test.norm.df1[, 1:14] <- predict(norm.values, Test_Data[, 1:14])
norm.values <- preProcess(Train_Data[, 1:14], method=c("center", "scale"))
```


```{r}
summary(train.norm.df1)
```


```{r}
summary(valid.norm.df1)
```


```{r}
##Let’s now run knn on the training set, and compare the confusion matrices on the validation and test sets
##Apply the k-NN method with the k = 3


 knn.pred <- knn(train.norm.df1[,-7], valid.norm.df1[,-7], 
                  cl = train.norm.df1[,7], k = 3)
accuracy.df <- confusionMatrix(knn.pred, valid.norm.df1[,7])$overall[1]
accuracy.df
  
```




```{r}
CrossTable(x=valid.norm.df1[,7],y=knn.pred)
```


```{r}
# Test Data
knn.pred <- knn(train.norm.df1[,-7], test.norm.df1[,-7], 
                  cl = train.norm.df1[,7], k = 3)
accuracy.df <- confusionMatrix(knn.pred, test.norm.df1[,7])$overall[1]
accuracy.df
```

```{r}
CrossTable(x=test.norm.df1[,7], y=knn.pred)
```

